InteroceptionScale (study 2) - Data Analysis

### 2025-07-02 study2/analysis/1_cleaning.qmd
update

dfchecks |>
  mutate(Total = as.factor(paste0(Total, "/8"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

@@

We removed `r length(outliers$attentionchecks)` (`r insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for having failed at least 1 attention check (out of 8).

>>>>

dfchecks |>
  mutate(Total = as.factor(paste0(Total, "/7"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

@@

We removed `r length(outliers$attentionchecks)` (`r insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for having failed at least 1 attention check (out of 7).

### 2025-07-02 study2/analysis/2_analysis.qmd
update

p_hclust <- tbl_graph(nodes = nodes, edges = edges) |> 
  ggraph(layout = "dendrogram", circular = TRUE) +
  # geom_edge_diagonal(strength = 0.7, linewidth = 1) +
  geom_edge_elbow2(aes(edge_width=linewidth), color="darkgrey") +
  geom_node_point(aes(filter=Group %in% c("Central"), size = Size), color="darkgrey") +
  geom_node_point(aes(filter=Group %in% c("Visceroception", "Awareness", "Deficit"), color=Group, size = Size)) +
  geom_node_text(aes(
    label = ifelse(label != "NA", label, NA),
    x = x * 1.05,
    y = y * 1.05,
    filter = label != "",
    angle = ifelse(
      x >= 0,
      asin(y) * 360 / 2 / pi,
      360 - asin(y) * 360 / 2 / pi
    ),
    hjust = ifelse(
      x >= 0, 0, 1
    ))) +
  geom_node_point(aes(filter = label != "", color=Group, size=Size), alpha = 1) +
  # geom_node_text(aes(label=idx)) +  # Debug
  scale_edge_width_continuous(range=c(1, 3), guide = "none") +
  scale_size_continuous(range=c(3, 11), guide = "none") +
  scale_color_manual(values = c(
    "Visceroception" = "#FF7811", "Card" = "#F44336", "Resp"="#FF9800", "Gast"="#795548",
    "Awareness" = "#4CAF50", "ExAc"="#4CAF50", "RelA"="#8BC34A", "SexS" = "#CDDC39",
    "Deficit" = "#9C27B0", "CaCo"="#9C27B0", "Urin" = "#673AB7", "Derm"="#3F51B5",
    "Sati" = "#2196F3", "Olfa" = "#00BCD4"),
                     breaks = c(
                       "Awareness", "ExAc", "RelA", "SexS",
                       "Deficit", "CaCo", "Urin", "Derm", "Sati", "Olfa",
                       "Visceroception", "Resp", "Card", "Gast"
                       ),
                     labels=c(
                       "***Awareness***", "ExAc", "RelA", "SexS",
                       "***Deficit***", "CaCo", "Urin", "Derm", "Sati", "Olfa",
                       "***Visceroception***", "Resp", "Card", "Gast"
                       )) +
  ggtitle("Hierarchical Clustering", subtitle = "Method = Correlation") +
  coord_equal(clip = "off", xlim = c(-1.25, 1.25), ylim = c(-1.25, 1.25)) +
  theme_void() + 
  guides(color = guide_legend(override.aes = list(size = c(7.5, 5, 5, 3.5, 7.5, 5, 5, 3.5, 3.5, 3.5, 7.5, 5, 5, 3.5)))) +
  theme(legend.text = ggtext::element_markdown(),
        legend.title = element_blank(),
        plot.title = element_blank(),  #element_text(face="bold")
        plot.subtitle = element_blank())
p_hclust
```

@@

ggsave("figures/fig1a.png", p_stab, width=10, height=10*0.8, dpi=300)
ggsave("figures/fig1b.png", p_hclust, width=12, height=8, dpi=300)
```

@@

  if(length(unique(df[[yvar]])) > 2) {
    # LMs - display correlation
    r_mint <- cor.test(cbind(df, metamint)[[best_pred_mint$Name[1]]], df[[yvar]])
    r_mint <- paste0(insight::format_value(r_mint$estimate, zap_small = TRUE))
    r_nonmint <- cor.test(df[[best_pred_nonmint$Name[1]]], df[[yvar]])
    r_nonmint <- paste0(insight::format_value(r_nonmint$estimate, zap_small = TRUE))
  } else {
    # GLMs - display log-odds
    r_mint <- best_pred_mint$Coefficient
    r_nonmint <- best_pred_nonmint$Coefficient
  }

@@

tab2 <- rez |> 
  gt::gt() |> 
  gt::tab_header(
    title = gt::md("**Interoceptive Questionnaires Comparison**"),
    subtitle = "MINT vs. MAIA, IAS, BPQ"
  ) |> 
  gt::cols_width(
    Outcome ~ gt::pct(23),
    Best_MINT_Predictor ~ gt::pct(12),
    Best_NonMINT_Predictor ~ gt::pct(15),
    Best_Models_R2 ~ gt::pct(25),
    Best_Models_BIC ~ gt::pct(25)
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight="bold", v_align = "middle"),
    locations = gt::cells_column_labels()
  )  |>
  gt::tab_style(
    style = gt::cell_text(align = "center", v_align = "middle"),
    locations = gt::cells_body(columns=c("Best_MINT_Predictor", "Best_NonMINT_Predictor"))
  )  |>
  gt::tab_row_group(
    label = "Lifestyle",
    rows = Group == "Lifestyle"
  ) |> 
  gt::tab_row_group(
    label = "Dissociative Symptoms",
    rows = Group == "Dissociative Symptoms"
  ) |> 
  gt::tab_row_group(
    label = "Primal World Beliefs",
    rows = Group == "Primals"
  ) |> 
  gt::tab_row_group(
    label = "Somatic Health",
    rows = Group == "Somatic Health"
  ) |> 
  gt::tab_row_group(
    label = "Mental Health",
    rows = Group == "Mental Health"
  ) |>
  gt::tab_row_group(
    label = "Mood",
    rows = Group == "Mood"
  ) |> 
  gt::tab_row_group(
    label = "Emotional Reactivity",
    rows = Group == "Emotion Reactivity"
  ) |> 
  gt::tab_row_group(
    label = "Alexithymia",
    rows = Group == "Alexithymia"
  ) |> 
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "gray85"),
      gt::cell_text(weight="bold", style="italic")),
    locations = gt::cells_row_groups()
  )  |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(1:4, 11:18, 20:21, 24))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(5:9, 10, 19, 22))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#EF9A9A"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(23, 25))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_Models_R2", rows=c(1:3, 11:13, 15:21, 23:25))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_Models_R2", rows=c(4:10, 14, 22))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(1:3, 11:20, 25))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(4:10, 22))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#EF9A9A"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(21, 23, 24))
    )
  ) |>
  gt::cols_label(
    Best_MINT_Predictor = "Best Predictor\n(MINT)",
    Best_NonMINT_Predictor = "Best Predictor\n(Non-MINT)",
    Best_Models_R2 = "Best Models (R2)",
    Best_Models_BIC = "Best Models (BIC)"
  ) |> 
  gt::cols_hide(Group) |> 
  gt::fmt_markdown() |> 
  gt::tab_options(
    table.font.size = gt::pct(125)
  )

@@

gt::gtsave(tab2, "figures/table2.png", vwidth=1400, expand=0, quiet=TRUE)
```

@@

fig1a <- patchwork::wrap_elements(
  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/fig1a.png"))) +
  patchwork::plot_annotation(title = "Exploratory Graphical Analysis (EGA)",
                             subtitle = "Bootstrapped replication of hierarchical clusters (Methhod = leiden)",
                             theme = theme(plot.title = element_text(face="bold", hjust = 0, size = rel(1.4)),
                                           plot.subtitle = element_text(hjust = 0, size = rel(1)))))
fig1b <- patchwork::wrap_elements(
  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/fig1b.png"))) +
  patchwork::plot_annotation(title = "Hierarchical Clustering",
                             subtitle = "Method = Correlation",
                             theme = theme(plot.title = element_text(face="bold", hjust = 0, size = rel(1.4)),
                                           plot.subtitle = element_text(hjust = 0, size = rel(1)))))
ggsave("figures/fig1.png", fig1a / fig1b, width=8, height=12, dpi=300)

>>>>

p_hclust <- tbl_graph(nodes = nodes, edges = edges) |> 
  ggraph(layout = "dendrogram", circular = TRUE) +
  # geom_edge_diagonal(strength = 0.7, linewidth = 1) +
  geom_edge_elbow2(aes(edge_width=linewidth), color="darkgrey") +
  geom_node_point(aes(filter=Group %in% c("Central"), size = Size), color="darkgrey") +
  geom_node_point(aes(filter=Group %in% c("Visceroception", "Awareness", "Deficit"), color=Group, size = Size)) +
  geom_node_text(aes(
    label = ifelse(label != "NA", label, NA),
    x = x * 1.07,
    y = y * 1.07,
    filter = label != "",
    angle = ifelse(
      x >= 0,
      asin(y) * 360 / 2 / pi,
      360 - asin(y) * 360 / 2 / pi
    ),
    hjust = ifelse(
      x >= 0, 0, 1
    ))) +
  geom_node_point(aes(filter = label != "", color=Group, size=Size), alpha = 1) +
  # geom_node_text(aes(label=idx)) +  # Debug
  scale_edge_width_continuous(range=c(1, 3), guide = "none") +
  scale_size_continuous(range=c(3, 11), guide = "none") +
  scale_color_manual(values = c(
    "Visceroception" = "#FF7811", "Card" = "#F44336", "Resp"="#FF9800", "Gast"="#795548",
    "Awareness" = "#4CAF50", "ExAc"="#4CAF50", "RelA"="#8BC34A", "SexS" = "#CDDC39",
    "Deficit" = "#9C27B0", "CaCo"="#9C27B0", "Urin" = "#673AB7", "Derm"="#3F51B5",
    "Sati" = "#2196F3", "Olfa" = "#00BCD4"),
                     breaks = c(
                       "Awareness", "ExAc", "RelA", "SexS",
                       "Deficit", "CaCo", "Urin", "Derm", "Sati", "Olfa",
                       "Visceroception", "Resp", "Card", "Gast"
                       ),
                     labels=c(
                       "***Awareness***", "ExAc", "RelA", "SexS",
                       "***Deficit***", "CaCo", "Urin", "Derm", "Sati", "Olfa",
                       "***Visceroception***", "Resp", "Card", "Gast"
                       )) +
  ggtitle("Hierarchical Clustering", subtitle = "Method = Correlation") +
  coord_equal(clip = "off", xlim = c(-1.25, 1.25), ylim = c(-1.25, 1.25)) +
  theme_void() + 
  guides(color = guide_legend(override.aes = list(size = c(7.5, 5, 5, 3.5, 7.5, 5, 5, 3.5, 3.5, 3.5, 7.5, 5, 5, 3.5)))) +
  theme(legend.text = ggtext::element_markdown(),
        legend.title = element_blank(),
        plot.title = element_blank(),  #element_text(face="bold")
        plot.subtitle = element_blank())
p_hclust
```

@@

ggsave("figures/fig1a.png", p_stab, width=10, height=10*0.8, dpi=300)
ggsave("figures/fig1b.png", p_hclust, width=8, height=8*0.8, dpi=300)
```

@@

  if(length(unique(df[[yvar]])) > 2) {
    # LMs - display correlation
    r_mint <- cor.test(cbind(df, metamint)[[best_pred_mint$Name[1]]], df[[yvar]])
    r_mint <- paste0(insight::format_value(r_mint$estimate, zap_small = TRUE))
    r_nonmint <- cor.test(df[[best_pred_nonmint$Name[1]]], df[[yvar]])
    r_nonmint <- paste0(insight::format_value(r_nonmint$estimate, zap_small = TRUE))
  } else {
    # GLMs - display log-odds
    r_mint <- insight::format_value(best_pred_mint$Coefficient)
    r_nonmint <- insight::format_value(best_pred_nonmint$Coefficient)
  }

@@

tab2 <- rez |> 
  gt::gt() |> 
  gt::tab_header(
    title = gt::md("**Interoception Questionnaires Comparison**"),
    subtitle = "MINT vs. MAIA, IAS, BPQ"
  ) |> 
  gt::cols_width(
    Outcome ~ gt::pct(23),
    Best_MINT_Predictor ~ gt::pct(12),
    Best_NonMINT_Predictor ~ gt::pct(15),
    Best_Models_R2 ~ gt::pct(25),
    Best_Models_BIC ~ gt::pct(25)
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight="bold", v_align = "middle"),
    locations = gt::cells_column_labels()
  )  |>
  gt::tab_style(
    style = gt::cell_text(align = "center", v_align = "middle"),
    locations = gt::cells_body(columns=c("Best_MINT_Predictor", "Best_NonMINT_Predictor"))
  )  |>
  gt::tab_row_group(
    label = "Lifestyle",
    rows = Group == "Lifestyle"
  ) |> 
  gt::tab_row_group(
    label = "Dissociative Symptoms",
    rows = Group == "Dissociative Symptoms"
  ) |> 
  gt::tab_row_group(
    label = "Primal World Beliefs",
    rows = Group == "Primals"
  ) |> 
  gt::tab_row_group(
    label = "Somatic Health",
    rows = Group == "Somatic Health"
  ) |> 
  gt::tab_row_group(
    label = "Mental Health",
    rows = Group == "Mental Health"
  ) |>
  gt::tab_row_group(
    label = "Mood",
    rows = Group == "Mood"
  ) |> 
  gt::tab_row_group(
    label = "Emotional Reactivity",
    rows = Group == "Emotion Reactivity"
  ) |> 
  gt::tab_row_group(
    label = "Alexithymia",
    rows = Group == "Alexithymia"
  ) |> 
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "gray85"),
      gt::cell_text(weight="bold", style="italic")),
    locations = gt::cells_row_groups()
  )  |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(1:4, 11:18, 20:21, 24))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(5:9, 10, 19, 22))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#EF9A9A"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(23, 25))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_Models_R2", rows=c(1:3, 11:13, 15:21, 23:25))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_Models_R2", rows=c(4:10, 14, 22))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(1:3, 11:20, 25))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(4:10, 22))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#EF9A9A"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(21, 23, 24))
    )
  ) |>
  gt::cols_label(
    Best_MINT_Predictor = "Best Predictor\n(MINT)",
    Best_NonMINT_Predictor = "Best Predictor\n(Non-MINT)",
    Best_Models_R2 = "Best Models (R2)",
    Best_Models_BIC = "Best Models (BIC)"
  ) |> 
  gt::cols_hide(Group) |> 
  gt::fmt_markdown() |> 
  gt::tab_options(
    table.font.size = gt::pct(125)
  )

@@

# gt::gtsave(tab2, "figures/table2.png", vwidth=1400, expand=0, quiet=TRUE)
# gt::gtsave(tab2, "figures/table2b.png", vwidth=1600, expand=0, quiet=TRUE)

tab2 |> 
  gt::cols_width(
    Outcome ~ gt::pct(21),
    Best_MINT_Predictor ~ gt::pct(17),
    Best_NonMINT_Predictor ~ gt::pct(18),
    Best_Models_R2 ~ gt::pct(22),
    Best_Models_BIC ~ gt::pct(22)
  ) |> 
  gt::gtsave("figures/table2.png", vwidth=1700, expand=0, quiet=TRUE)
```

@@

fig1a <- patchwork::wrap_elements(
  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/fig1a.png"))) +
  patchwork::plot_annotation(title = "Exploratory Graphical Analysis (EGA)",
                             subtitle = "Bootstrapped replication of clusters (Method = Leiden)",
                             theme = theme(plot.title = element_text(face="bold", hjust = 0, size = rel(1.4)),
                                           plot.subtitle = element_text(hjust = 0, size = rel(1)))))
fig1b <- patchwork::wrap_elements(
  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/fig1b.png"))) +
  patchwork::plot_annotation(title = "Hierarchical Clustering",
                             subtitle = "Method = Correlation",
                             theme = theme(plot.title = element_text(face="bold", hjust = 0, size = rel(1.4)),
                                           plot.subtitle = element_text(hjust = 0, size = rel(1)))))
ggsave("figures/fig1.png", fig1a / fig1b, width=8, height=12, dpi=300)
